Reproduction frequency and offspring survival decline in elephant seals past prime age

### 2024-10-10 paper/paperv3.qmd
Adding a comparison to a linear model

# Are hypotheses supported by the model? note to future readers...this part could have been done way easier, and you will have to trust our process that this was the path of least resistence for us at the point of final revisions!
extract_p <- function(h, tails = 1) {
  if (tails == 1) {
    map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
  } else if (tails == 2) {
    map_dbl(mod_sens, \(m) 2 * pnorm(m[[h]]["age10:age_catOld", "z value"],
                                     lower.tail = (m[[h]]["age10:age_catOld", "z value"] < 0)
                                     ))
  } else {
    stop("tails must be 1 or 2")
  }
}
extract_aic <- function(h) {
  map_dbl(mod_sens, paste(h, "aic", sep = "_"))
}
p_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_p("h1", tails = 2),
                "H2a (survival)" = extract_p("h2_surv", tails = 2),
                "H2b (recruitment)" = extract_p("h2_recr", tails = 2),
                "H3 (sex ratio; 2 tail)" = extract_p("h3", tails = 2))
aic_weight <- function(x) {
  rellik <- exp(-0.5 * (x - min(x)))
  rellik / sum(rellik)
}
aic_tbl <- tibble(age_cutoff = cutoff,
                  "H1 (breeding)" = extract_aic("h1"),
                  "H2a (survival)" = extract_aic("h2_surv"),
                  "H2b (recruitment)" = extract_aic("h2_recr"),
                  "H3 (sex ratio)" = extract_aic("h3")) %>% 
  mutate(across(-age_cutoff, aic_weight))

@@

## Breakpoint vs quadratic

@@

@@

@@

aic <- function(break_mod, quad_mod, hyp) {
  AIC(break_mod, quad_mod) %>% 
    as_tibble() %>% 
    mutate(dAIC = AIC - min(AIC),
         rellik = exp(-0.5 * dAIC),
         weight = rellik / sum(rellik),
         hypothesis = hyp,
         model = c("Breakpoint", "Quadratic")) %>% 
  select(hypothesis, model, AIC, weight)
}

pmap(list(break_mod = list(breed_age_mod, pup_surv_mod, pup_recr_mod),
          quad_mod = list(breed_age_quad, surv_age_quad, recr_age_quad),
          hyp = c("H1", "H2a", "H2b")),
     aic) %>% 
  list_rbind() %>% 
  knitr::kable()
```

>>>>

# Are hypotheses supported by the model? note to future readers...this part could have been done way easier, and you will have to trust our process that this was the path of least resistence for us at the point of final revisions!
extract_p <- function(h, tails = 1) {
  if (tails == 1) {
    map_dbl(mod_sens, \(m) pnorm(m[[h]]["age10:age_catOld", "z value"]))
  } else if (tails == 2) {
    map_dbl(mod_sens, \(m) 2 * pnorm(m[[h]]["age10:age_catOld", "z value"],
                                     lower.tail = (m[[h]]["age10:age_catOld", "z value"] < 0)
                                     ))
  } else {
    stop("tails must be 1 or 2")
  }
}
extract_aic <- function(h) {
  map_dbl(mod_sens, paste(h, "aic", sep = "_"))
}
p_tbl <- tibble(age_cutoff = cutoff, 
                "H1 (breeding)" = extract_p("h1", tails = 2),
                "H2a (survival)" = extract_p("h2_surv", tails = 2),
                "H2b (recruitment)" = extract_p("h2_recr", tails = 2),
                "H3 (sex ratio)" = extract_p("h3", tails = 2))
aic_weight <- function(x) {
  rellik <- exp(-0.5 * (x - min(x)))
  rellik / sum(rellik)
}
aic_tbl <- tibble(age_cutoff = cutoff,
                  "H1 (breeding)" = extract_aic("h1"),
                  "H2a (survival)" = extract_aic("h2_surv"),
                  "H2b (recruitment)" = extract_aic("h2_recr"),
                  "H3 (sex ratio)" = extract_aic("h3")) %>% 
  mutate(across(-age_cutoff, aic_weight))

@@

## Breakpoint vs quadratic vs linear

@@

```{r}
#| label: linear-h1

breed_age_line <- lmer(
  observed_int ~ age10 + (1 | animalID) + (1 | year),
  sealdat
)

summary(breed_age_line)$coefficients
```

@@

```{r}
#| label: linear-h2

#h2a
surv_age_line <- lmer(
  pup_surv_int ~ age10 + (1 | animalID) + (1 | year),
  surv_recr_data
)

summary(surv_age_line)$coefficients

#h2b
recr_age_line <- lmer(
  pup_recr_int ~ age10 + (1 | year),
  filter(surv_recr_data, pupsex == "F")
)

summary(recr_age_line)$coefficients
```

@@

aic <- function(break_mod, quad_mod, line_mod, hyp) {
  AIC(break_mod, quad_mod, line_mod) %>% 
    as_tibble() %>% 
    mutate(dAIC = AIC - min(AIC),
         rellik = exp(-0.5 * dAIC),
         weight = rellik / sum(rellik),
         hypothesis = hyp,
         model = c("Breakpoint", "Quadratic", "Linear")) %>% 
  select(hypothesis, model, AIC, weight)
}

pmap(list(break_mod = list(breed_age_mod, pup_surv_mod, pup_recr_mod),
          quad_mod = list(breed_age_quad, surv_age_quad, recr_age_quad),
          line_mod = list(breed_age_line, surv_age_line, recr_age_line), 
          hyp = c("H1", "H2a", "H2b")),
     aic) %>% 
  list_rbind() %>% 
  knitr::kable()
```
